---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
```{r}
library(dplyr)
library(lattice)

mydata <- read.csv("./activity.csv")
mydata$date<-as.Date(mydata$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r}
ActivityByDay <- group_by(mydata,date)
TotalByDay <- summarize(ActivityByDay,sum(steps,na.rm=TRUE))

hist(TotalByDay$`sum(steps, na.rm = TRUE)`,
     main= "Frequency of Days by Number of Steps",
     xlab = "Steps on any one Day",
     ylab = "Number of Days",
     breaks=20)

DailyStepMean <-mean(TotalByDay$`sum(steps, na.rm = TRUE)`)
DailyStepMedian <-median(TotalByDay$`sum(steps, na.rm = TRUE)`)

```
The mean number of steps per day is `r DailyStepMean`.  
The median number of steps per day is `r DailyStepMedian`.

## What is the average daily activity pattern?

```{r}
meanbyint <- summarize(group_by(mydata,interval),mean(steps,na.rm = TRUE))

plot(meanbyint,
      type = "l",
      main = "Average Steps by Interval",
      xlab = "5 Minute Interval",
      ylab = "Average Steps")
```

What time interval has the maximum average steps?

```{r}

highestint<- subset(meanbyint,meanbyint$`mean(steps, na.rm = TRUE)`==max(meanbyint$`mean(steps, na.rm = TRUE)`))

```

The interval with the highest number of steps is `r highestint[1,1]`.  

## Imputing missing values

Some rows have NA Values.
How many rows have NA values?
```{r}
NArows <- sum(is.na(mydata$steps))

```

The NA values will be imputed using the average for a particular 5 minute interval

```{r}
mydata_na<-subset(mydata,is.na(mydata$steps))
mydata_pop<-subset(mydata,is.na(mydata$steps)==FALSE)

mydata_na_mean <- left_join(mydata_na,meanbyint,by="interval")
mydata_na_mean$steps=NULL
names(mydata_na_mean)[3]<-"steps"

mydata_na_impute <- rbind(mydata_na_mean,mydata_pop)
```
## What is mean total number of steps taken per day using imputed values?

```{r}
ActivityByDay_na_impute <- group_by(mydata_na_impute,date)
TotalByDay_na_impute <- summarize(ActivityByDay_na_impute,sum(steps,na.rm=TRUE))

hist(TotalByDay_na_impute$`sum(steps, na.rm = TRUE)`,
     main= "Frequency of Days by Number of Steps",
     xlab = "Steps on any one Day",
     ylab = "Number of Days",
     breaks=20)

DailyStepMean2 <-mean(TotalByDay_na_impute$`sum(steps, na.rm = TRUE)`)
DailyStepMedian2 <-median(TotalByDay_na_impute$`sum(steps, na.rm = TRUE)`)

```
The mean number of steps per day is `r DailyStepMean2`.  
The median number of steps per day is `r DailyStepMedian2`.

It appears that imputing the NA values increases the mean by `r DailyStepMean2 - DailyStepMean` days.  
It appears that imputing the NA values increases the median by `r DailyStepMedian2 - DailyStepMedian` days.  

## Are there differences in activity patterns between weekdays and weekends?

First add a flag for the weekend
```{r}
mydata$weekday <- weekdays(mydata$date)
mydata$Type<-"Weekday"

mydata$Type[mydata$weekday=="Saturday"]<-"Weekend"
mydata$Type[mydata$weekday=="Sunday"]<-"Weekend"

```

Next graph the data

```{r}
meanbyint_wknd <- summarize(group_by(mydata,interval,Type),mean(steps,na.rm = TRUE))

xyplot(meanbyint_wknd$`mean(steps, na.rm = TRUE)`~meanbyint_wknd$interval|meanbyint_wknd$Type,
       layout=c(1,2),
       type="l",
       main="Average Steps by Interval - Weekend vs. Weekday",
       xlab = "Interval",
       ylab="Average Steps")

```